---
title: "Interactive plots"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  cache = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%"
)
```

To illustrate how interactive plots work in **ebm**, we'll fit a simple EBM regression model to the `Hitters` data from the [ISLR2](https://cran.r-project.org/package=ISLR2) package. 

```{r data}
library(ebm)

# Load the data and remove rows with missing response values
data("Hitters", package = "ISLR2")
hitters <- Hitters[!is.na(Hitters$Salary), ]

# Fit a simple EBM; here we turn off some of the more expensive arguments
fit <- ebm(Salary ~ ., data = hitters, objective = "rmse", inner_bags = 0,
           outer_bags = 1, interactions = 0)
```

You can produce several [plotly](https://plotly.com/r/)-based graphs to help interpret the output of `"EBM"` objects by calling the generic `plot()` method with `interactive = TRUE`; this function supports both global and local interpretations. The default simply prints a global measure of importance based on the mean absolute score for each term. (For Markdown-type documents, like this article, you also need to specify `display = "markdown"`; see `?ebm::plot` for details.)

```{r ebm-regressor-show-global-importance}
# Plot feature importance (i.e., mean absolute scores)
plot(fit, interactive = TRUE, display = "markdown")
```

You can also plot the individual shape functions (or term contributions), as shown below:

```{r ebm-regressor-show-global-term}
plot(fit, term = "Years", interactive = TRUE, display = "markdown")
```

You can also display local explanations (though, one at a time) by specifying `local = TRUE`:

```{r ebm-regressor-show-local-interactive}
# Understand an individual prediction
x <- subset(hitters, select = -Salary)[1L, ]  # use first observation
plot(fit, local = TRUE, X = x, y = hitters$Salary[1L], 
     interactive = TRUE, display = "markdown")
```

